Boosted Mirror Descent, Frank-Wolfe, and Herding

### 2013-09-07 icml/paper.tex
made few small edits and added a couple comments

Consider the following loss function:
\[ L(u) = \max_{\theta \in \Theta} [\langle \theta, u \rangle - R(\theta)] \]
$R$ is some convex function that we should think of as a measure of complexity 
of $\theta$. Thus, the loss is equal to the maximum amount that $\theta$ can match 
$u$, regularized by the complexity function $R(\theta)$.

Examples:
\begin{itemize}
\item Let $R(\theta) = \langle \theta, u_0 \rangle$. Then $L(u)$ is the 
      \emph{maximum mean discrepancy} between $u$ and $u_0$ 
      relative to $\Theta$.
\item Let $R(\theta) = S^*(\theta)$ where $S^*$ is the Fenchel conjugate of 
      any strongly convex function $S$. Then $L(u) = S(u)$.
\end{itemize}

@@

First, let us generalize the setting 
to consider a \emph{two-argument} loss function
\[ L(u,\theta) = h(u) + \langle \theta, u \rangle - R(\theta). \]
In the case $h(u) \equiv 0$, we can get the previous setting by 
considering $\max_{\theta \in \Theta} L(u,\theta)$. We will assume 
throughout this document that $h$ and $R$ are both convex functions.

Boosted mirror descent can be thought of as an algorithm for finding 
``saddle points'' of $L$. It is the following algorithm:
\begin{enumerate}
\item $u_1 = 0$
\item $\theta_{t} \in \arg\max_{\theta \in \Theta} \langle \theta, u_t \rangle - R(\theta)$
\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \frac{1}{t} \sum_{s \leq t} \theta_s, u \rangle$
\end{enumerate}
Let $\hat{u} = \frac{1}{T} \sum_{t = 1}^{T} u_t$. As long as $h$ is strongly convex, we obtain the 
bound:
\[ \sup_{\theta \in \Theta} L(\hat{u}, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(1/\sqrt{T}). \]
However, it is often the case that $h$ is not strongly convex, whereas $R$ is strongly convex. In this case 
we may wish to use the following slightly different algorithm:
\begin{enumerate}
\item $\theta_1 = 0$
\item $u_t \in \arg\min_{u} h(u) + \langle \theta_t, u \rangle$
\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, \frac{1}{t} \sum_{s \leq t} u_s \rangle - R(\theta)$
\end{enumerate}
Once again, if we let $\hat{u} = \frac{1}{T} \sum_{t=1}^T u_t$, then we obtain the following bound, which now 
only requires the convexity of $R$:
\[ \sup_{\theta \in \Theta} L(\hat{u}, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(1/\sqrt{T}). \]

@@

\section{Application to Herding}
\label{sec:herding}
Suppose we are given a family of features $\phi$ together 
with a known value $\bar{\phi}$ for $\bE_{\mu}[\phi]$. We 
would like to construct a distribution that approximately 
matches this distribution, i.e. one for which 
$\bE_{\hat{\mu}}[\phi] \approx \bar{\phi}$. A natural way 
to do this is by using boosted mirror descent to minimize the 
maximum mean discrepancy relative to $\phi$; 
in other words, to minimize
\[ L(\mu) = \max\limits_{i=1}^d |\bE_{\mu}[\phi_i]-\bar{\phi}_i|. \]
If we wish to write this more similarly to the original problem 
formulation, we can write
\[ L(\mu) = \sup_{\|\theta\|_1 \leq 1} \theta^T(\bE_{\mu}[\phi]-\bar{\phi}). \]
One issue is that $\|\theta\|_1$ is not strongly convex, so we relax $\|\theta\|_1$ 
to $\|\theta\|_2$ and write our loss function in the Lagrangian form:
\[ L(\mu, \theta) = \theta^T\bE_{\mu}[\phi] - \theta^T\bar{\phi} - \frac{1}{2}\|\theta\|_2^2. \]
Intriguingly, Method 1 and Method 2 are identical to each other for this choice of $L$ 
(up to a change of index). In this case, $\mu_t = \delta_{x_t}$ for some $x_t$, and 
the updates are the same as the standard herding updates. Let $R = \sup_{x} \|\phi(x)\|_2$; 
then using the Method 2 convergence bound gives us
\begin{align*}
\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T \delta_{x_t}, \theta\right) &\leq \inf_{u^*} \sup_{\theta} L(u^*, \theta) + \frac{2R^2\log(T+1)}{T} \\
 &= \sup_{\theta} \inf_{u^*} L(u^*, \theta) + \frac{2R^2\log(T+1)}{T} \\
 &= \frac{2R^2\log(T+1)}{T}.
\end{align*}
This is a slightly weaker version of the typical herding bound on MMD.

>>>>

Consider the following loss function:
\[ L(u) = \max_{\theta \in \Theta} \left\{\langle \theta, u \rangle - R(\theta) \right\}, \]
where $R$ is some convex function that we should think of as a measure of complexity 
of $\theta$. Thus, the loss is equal to the maximum amount that $\theta$ can match 
$u$, regularized by the complexity function $R(\theta)$.

Examples:
\begin{itemize}
\item Let $R(\theta) = \langle \theta, u_0 \rangle$. Then, assuming $-\Theta = \Theta$, $L(u)$ is the 
      \emph{maximum mean discrepancy} between $u$ and $u_0$ 
      relative to $\Theta$.
\item Let $R(\theta) = S^*(\theta)$ where $S^*$ is the Fenchel conjugate of 
      any strongly convex function $S$. Then $L(u) = S(u)$.
\end{itemize}

@@

First, let us generalize the setting 
to consider a \emph{two-argument} loss function
\[ L(u,\theta) = h(u) + \langle \theta, u \rangle - R(\theta). \]
In the case $h(u) \equiv 0$, we can get the previous setting by 
considering $\max_{\theta \in \Theta} L(u,\theta)$. We will assume 
throughout this document that $h$ and $R$ are both convex functions. \NA{JH: Should we comment on the symmetry introduced by this formulation?}

Boosted mirror descent can be thought of as an algorithm for finding 
``saddle points'' of $L$. It is the following algorithm:
\begin{enumerate}
\item $u_1 = 0$
\item $\theta_{t} \in \arg\max_{\theta \in \Theta} \langle \theta, u_t \rangle - R(\theta)$
\item $u_{t+1} \in \arg\min_{u} h(u) + \langle \frac{1}{t} \sum_{s \leq t} \theta_s, u \rangle$
\end{enumerate}
Let $\hat{u} = \frac{1}{T} \sum_{t = 1}^{T} u_t$. As long as $h$ is strongly convex, we obtain the 
bound:
\[ \sup_{\theta \in \Theta} L(\hat{u}, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(\log(T)/T). \]
However, it is often the case that $h$ is not strongly convex, whereas $R$ is strongly convex. In this case 
we may wish to use the following slightly different algorithm:
\begin{enumerate}
\item $\theta_1 = 0$
\item $u_t \in \arg\min_{u} h(u) + \langle \theta_t, u \rangle$
\item $\theta_{t+1} \in \arg\max_{\theta \in \Theta} \langle \theta, \frac{1}{t} \sum_{s \leq t} u_s \rangle - R(\theta)$
\end{enumerate}
Once again, if we let $\hat{u} = \frac{1}{T} \sum_{t=1}^T u_t$, then we obtain the following bound, which now 
only requires the convexity of $R$:
\[ \sup_{\theta \in \Theta} L(\hat{u}, \theta) \leq \sup_{\theta \in \Theta} L(u^*, \theta) + O(\log(T)/T). \]
\NA{JH: It would be more elegant to argue either here or below that the bounds on the second algorithm follow directly from duality}

@@

\section{Application to Herding}
\label{sec:herding}
Suppose we are given a family of features $\phi$ together 
with a known value $\bar{\phi}$ for $\bE_{\mu}[\phi]$. We 
would like to construct a distribution that approximately 
matches this distribution, i.e.~one for which 
$\bE_{\hat{\mu}}[\phi] \approx \bar{\phi}$. A natural way 
to do this is by using boosted mirror descent to minimize the 
maximum mean discrepancy relative to $\phi$; 
in other words, to minimize
\[ L(\mu) = \max_{i \in \{1,\dots,d\}} |\bE_{\mu}[\phi_i]-\bar{\phi}_i|. \]
If we wish to write this more similarly to the original problem 
formulation, we can write
\[ L(\mu) = \sup_{\|\theta\|_1 \leq 1} \theta^T(\bE_{\mu}[\phi]-\bar{\phi}). \]
One issue is that $\|\theta\|_1$ is not strongly convex, so we relax $\|\theta\|_1$ 
to $\|\theta\|_2$ and write our loss function in the Lagrangian form:
\[ L(\mu, \theta) = \theta^T\bE_{\mu}[\phi] - \theta^T\bar{\phi} - \frac{1}{2}\|\theta\|_2^2. \]
Intriguingly, Method 1 and Method 2 are identical to each other for this choice of $L$ 
(up to a change of index). In this case, $\mu_t = \delta_{x_t}$ for some $x_t$, and 
the updates are the same as the standard herding updates. Let $R = \sup_{x} \|\phi(x)\|_2$; 
then using the Method 2 convergence bound gives us
\begin{align*}
\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T \delta_{x_t}, \theta\right) &\leq \inf_{u^*} \sup_{\theta} L(u^*, \theta) + \frac{2R^2\log(T+1)}{T} \\
 &= \sup_{\theta} \inf_{u^*} L(u^*, \theta) + \frac{2R^2\log(T+1)}{T} \\
 &= \frac{2R^2\log(T+1)}{T}.
\end{align*}
This is a slightly weaker version of the typical herding bound on MMD.
